Estimation of Spatiotemporal Integro-Difference Equations for Dynamical Neural Field Models

### 2009-11-19 ltx/BrainIDE.tex
general edits
Alternative model- Unlimited function for mean pulse rate

\begin{figure*}
\centering
\includegraphics[scale=0.3]{ExampleFieldDecomposition.eps}
\caption[ExampleFieldDecomposition]{Illustration of a one dimensional field decomposition using 5 Gaussian basis functions, each weighted an element of the state vector. Each basis function has a width of 0.01 and is uniformly positioned with a separation of 0.25 (with arbitrary units).}
\label{ExampleFieldDecomposition}
\end{figure*}

@@

\section{State-Space Model}
If we are going to use a particle filter or a extended Kalman filter we might as well go for this case. If the temporal and spatial part of the nonlinearity can not be separated, than this is the best choice.
\begin{equation}\label{SystemEq}
{v_{t + 1}}\left( {\bf{r}} \right) = \int_\Omega  {w\left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}}  - \alpha {v_t}\left( {\bf{r}} \right) + {\varepsilon _t}\left( {\bf{r}} \right)
\end{equation}
Defining
\begin{equation}\label{CombiningKernels}
k\left( {{\bf{r}} - {\bf{r'}}} \right) = w\left( {{\bf{r}} - {\bf{r'}}} \right) - \hat{\alpha}\delta \left( {{\bf{r}} - {\bf{r'}}} \right).
\end{equation}
Making substitution we get
\begin{equation}\label{OneKernelModel}
{v_{t + 1}}\left( {\bf{r}} \right) = \int_\Omega  {k\left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}}  + {\varepsilon _t}\left( {\bf{r}} \right).
\end{equation}
Defining
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}
\end{equation}
\begin{equation}\label{DefKernelDecomp}
k\left( {{\bf{r}} - {\bf{r'}}} \right) = {\theta ^T}\psi \left( {{\bf{r}} - {\bf{r'}}} \right)
\end{equation}
Now we have four basis functions (the fourth being the delta/local connectivity) and four parameters to estimate. Making substitutions
\begin{equation}\label{DecompModel}
{\phi ^T}\left( {\bf{r}} \right){x_t} = {\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}}  + {\varepsilon _t}\left( {\bf{r}} \right)
\end{equation}
Cross multiply and integrate over $\bf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_t}}  = \int_\Omega  {\phi \left( {\bf{r}} \right){\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  + \int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
Defining
\begin{equation}\label{DefPsi}
\Psi  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
Making substitution
\begin{equation}\label{SS Model}
{x_t} = {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  + {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
\begin{equation}\label{AbbrevSSModel}
{x_{t + 1}} = f\left( {x;\theta } \right) + {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}

\section{Stability Analysis}
This section will not probably be included

>>>>

% \begin{figure*}
% \centering
% \includegraphics[scale=0.3]{ExampleFieldDecomposition.eps}
% \caption[ExampleFieldDecomposition]{Illustration of a one dimensional field decomposition using 5 Gaussian basis functions, each weighted an element of the state vector. Each basis function has a width of 0.01 and is uniformly positioned with a separation of 0.25 (with arbitrary units).}
% \label{ExampleFieldDecomposition}
% \end{figure*}

@@

\section{State-Space Model}
If we are going to use a particle filter or a extended Kalman filter we might as well go for this case. If the temporal and spatial part of the nonlinearity can not be separated, than this is the best choice.
\begin{equation}\label{SystemEq}
{v_{t + 1}}\left( {\bf{r}} \right) = \int_\Omega  {w\left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}}  - \alpha {v_t}\left( {\bf{r}} \right) + {\varepsilon _t}\left( {\bf{r}} \right)
\end{equation}
Defining
\begin{equation}\label{CombiningKernels}
k\left( {{\bf{r}} - {\bf{r'}}} \right) = w\left( {{\bf{r}} - {\bf{r'}}} \right) - \hat{\alpha}\delta \left( {{\bf{r}} - {\bf{r'}}} \right).
\end{equation}
Making substitution we get
\begin{equation}\label{OneKernelModel}
{v_{t + 1}}\left( {\bf{r}} \right) = \int_\Omega  {k\left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{v_t}\left( {{\bf{r'}}} \right)} \right)d{\bf{r'}}}  + {\varepsilon _t}\left( {\bf{r}} \right).
\end{equation}
Defining
\begin{equation}\label{DefFieldDecomp}
{v_t}\left( {\bf{r}} \right) = {\phi ^T}\left( {\bf{r}} \right){x_t}
\end{equation}
\begin{equation}\label{DefKernelDecomp}
k\left( {{\bf{r}} - {\bf{r'}}} \right) = {\theta ^T}\psi \left( {{\bf{r}} - {\bf{r'}}} \right)
\end{equation}
Now we have four basis functions (the fourth being the delta/local connectivity) and four parameters to estimate. Each connectivity basis function is equivalent to a layer in the WC model. Making substitutions
\begin{equation}\label{DecompModel}
{\phi ^T}\left( {\bf{r}} \right){x_{t+1}} = {\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}}  + {\varepsilon _t}\left( {\bf{r}} \right)
\end{equation}
Cross multiply and integrate over $\bf{r}$
\begin{equation}\label{Poo}
\int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}{x_{t+1}}}  = \int_\Omega  {\phi \left( {\bf{r}} \right){\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  + \int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
Defining
\begin{equation}\label{DefPsi}
\Psi  = \int_\Omega  {\phi \left( {\bf{r}} \right){\phi ^T}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
Making substitution
\begin{equation}\label{SS Model}
{x_{t+1}} = {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\theta ^T}\int_\Omega  {\psi \left( {{\bf{r}} - {\bf{r'}}} \right)p\left( {{\phi ^T}\left( {{\bf{r'}}} \right){x_t}} \right)d{\bf{r'}}} d{\bf{r}}}  + {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
\begin{equation}\label{AbbrevSSModel}
{x_{t + 1}} = f\left( {x;\theta } \right) + {\Psi ^{ - 1}}\int_\Omega  {\phi \left( {\bf{r}} \right){\varepsilon _t}\left( {\bf{r}} \right)d{\bf{r}}}
\end{equation}
\section{Alternative model-Unlimited function for $p(v_t(r))$}
We assume
\begin{equation}
 p(v_t(r))=\beta(v_t(\mathbf r)-v_0))=\beta v'_t(\mathbf r)
\end{equation}
adding $-v_0$ to the both side of the equation we have
\begin{equation}
 v_{t+1}(\mathbf s)-v_0=\beta\int_{\Omega} W(\mathbf s-\mathbf r) v'_t(\mathbf r)d\mathbf r-\frac{1}{\tau} v_t(\mathbf s)-v_0+e_t(\mathbf s)
\end{equation}
we add $\frac{1}{\tau}v_0-\frac{1}{\tau}v_0$ to the right side of the equation
\begin{equation}
 v'_{t+1}(\mathbf s)=\beta\int_{\Omega} W(\mathbf s-\mathbf r) v'_t(\mathbf r)d\mathbf r-\frac{1}{\tau} v'_t(\mathbf s)-(1+\frac{1}{\tau})v_0+e_t(\mathbf s)
\end{equation}

\begin{equation}
 v'_t(\mathbf s)=\boldsymbol \phi^{\top}(\mathbf s){\mathbf x}_t
\end{equation}
\begin{equation}
 w(\mathbf s -\mathbf r)=\boldsymbol \theta^{\top}\psi(\mathbf s -\mathbf r)
\end{equation}
\begin{equation}
\boldsymbol \phi^{\top}(\mathbf s){\mathbf x}_{t+1}=\beta\boldsymbol \theta^{\top}\boldsymbol\Phi(\mathbf s)\mathbf x_t-\frac{1}{\tau×}\boldsymbol \phi^{\top}(\mathbf s){\mathbf x}_t-(1+\frac{1}{\tau})v_0+e_t(\mathbf s)
\end{equation}
where
\begin{equation}
 \boldsymbol\Phi(\mathbf s)=\int_\Omega \psi(\mathbf s -\mathbf r)\boldsymbol \phi^{\top}(\mathbf r)d\mathbf r
\end{equation}
Cross multiplying by $\phi(s)$ and integrating over $\mathbf s$ gives
\begin{equation}
\int_{\Omega}\boldsymbol \phi(\mathbf s)\boldsymbol \phi^{\top}(\mathbf s)d\mathbf s{\mathbf x}_{t+1}=\beta\int_{\Omega}\boldsymbol \phi(\mathbf s) \theta^{\top}\boldsymbol\Phi(\mathbf s)d\mathbf s\mathbf x_t-\frac{1}{\tau×}\int_{\Omega}\boldsymbol \phi(\mathbf s)\boldsymbol \phi^{\top}(\mathbf s)d\mathbf s{\mathbf x}_t-(1+\frac{1}{\tau})v_0\int_{\Omega}\boldsymbol \phi(\mathbf s)d\mathbf s+\int_{\Omega}\boldsymbol \phi(\mathbf s)e_t(\mathbf s)d\mathbf s
\end{equation}
we define
\begin{equation}
 \boldsymbol\psi(\mathbf s)=\int_{\Omega}\boldsymbol \phi(\mathbf s)\boldsymbol \phi^{\top}(\mathbf s)d\mathbf s
\end{equation}
and
\begin{equation}
 \Gamma(\mathbf s)=\int_{\Omega}\boldsymbol \phi(\mathbf s) \theta^{\top}\boldsymbol\Phi(\mathbf s)d\mathbf s
\end{equation}
therefore
\begin{equation}
\boldsymbol\psi(\mathbf s){\mathbf x}_{t+1}=\beta\Gamma(\mathbf s)\mathbf x_t-\frac{1}{\tau×}\boldsymbol\psi(\mathbf s){\mathbf x}_t-(1+\frac{1}{\tau})v_0\int_{\Omega}\boldsymbol \phi(\mathbf s)d\mathbf s+\int_{\Omega}\boldsymbol \phi(\mathbf s)e_t(\mathbf s)d\mathbf s
\end{equation}
Multiplying by $\psi(\mathbf s)^{-1}$
\begin{equation}
 {\mathbf x}_{t+1}=\beta\psi(\mathbf s)^{-1}\Gamma(\mathbf s)\mathbf x_t-\frac{1}{\tau×}{\mathbf x}_t-(1+\frac{1}{\tau})v_0\psi(\mathbf s)^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf s)d\mathbf s+\psi(\mathbf s)^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf s)e_t(\mathbf s)d\mathbf s
\end{equation}
therefore
\begin{equation}
 {\mathbf x}_{t+1}=(\beta\psi(\mathbf s)^{-1}\Gamma(\mathbf s)-\frac{1}{\tau×}\mathbf I){\mathbf x}_t-\frac{1}{\tau×}{\mathbf x}_t-(1+\frac{1}{\tau})v_0\psi(\mathbf s)^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf s)d\mathbf s+\psi(\mathbf s)^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf s)e_t(\mathbf s)d\mathbf s
\end{equation}
\begin{equation}
 {\mathbf x}_{t+1}=\mathbf A(\boldsymbol\theta){\mathbf x}_t-(1+\frac{1}{\tau})v_0\psi(\mathbf s)^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf s)d\mathbf s+\psi(\mathbf s)^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf s)e_t(\mathbf s)d\mathbf s
\end{equation}
where
\begin{equation}
\mathbf A(\boldsymbol\theta)=(\beta\psi(\mathbf s)^{-1}\Gamma(\mathbf s)-\frac{1}{\tau×}\mathbf I)
\end{equation}
\section{Stability Analysis}
This section will not probably be included
